Douglas-complex fire smoke and cardiopulmonary morbidity

### 2017-11-16 r_scripts/analysis/oregon_cvd_pulm_results.Rmd
Fixed bugs and able to knit results document.

Packages used: tidyverse, survival.
```{r setup}
library(tidyverse) # general data wrangle
library(survival) # conditional logistic regression
```

@@

```{r smoke data import and wrangle}
# pm path
pm_path <- paste0("./data/pm/2013-oregon_county_pm25.csv")
# county pm
county_pm_df <- read_csv(pm_path) %>% 
  mutate(county = stringr::str_replace(county, "[.]", ""))

@@

```{r health data import, message=F, warning=F}
# path for health dataframe
health_path <- "./data/health/"
# importing time-stratified case-crossover dataframe
casecross_files <- list.files(path=health_path, pattern="*casecross*")

@@

```{r descriptive table function}
# creating a row calculation function
table1_fun <- function(data_list){
  # output of total n
  total_n <- data_list %>% 
    filter(outcome == 1) %>%
    # adding in a filter to remove NAs
    filter(!is.na(geo_smk_pm_zip)) %>% 
    summarise(total_n = n()) %>%

@@

# apply custom function to generate descriptive table
table_vals <- outcome_df_list %>% 
  # map to dataframe
  map_dfr(table1_fun) 
# descriptive table, bind outcome names and values
descriptive_table <- bind_cols(outcome, table_vals)

@@

```{r distributed lag}
# this approach will work that creates a list of dataframes with map then
# binds the lists together in one dataframe
dl_results <- outcome_df_list %>% 
  purrr::map(~distributed_lag(., pm_df=3, temp_df=2, cumulative=F)) %>% 
  plyr::rbind.fill() %>% 
  bind_cols(., data_frame(rep(outcome_names, each=8))) %>% 
  rename(outcome_names = `rep(outcome_names, each = 8)`) %>% 
  select(outcome_names, type, time, estimate, lower95, upper95) %>% 
  # preserve order of names
  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))

@@

```{r same day association stratified by sex}
# there is probably a better way to do this, but this will work.
sameday_results_f <- outcome_df_list %>% 
  # using gender because I forget which sex_ind is which
  # filter to females
  purrr::map(~filter(., gender == "F")) %>%  
  map_dfr(~same_day_fun(.)) %>% 
  bind_cols(., data_frame(outcome_names)) %>% 
  mutate(sex = as.factor("Female")) %>% 
  select(outcome_names, sex, estimates, lower95, upper95) %>% 
  # preserve order of names
  mutate(outcome_names = parse_factor(outcome_names, levels = outcome))

sameday_results_m <- outcome_df_list %>% 
  # filter to males
  purrr::map(~filter(., gender == "M")) %>%  
  map_dfr(~same_day_fun(.)) %>% 
  bind_cols(., data_frame(outcome_names)) %>% 
  mutate(sex = as.factor("Male")) %>% 
  select(outcome_names, sex, estimates, lower95, upper95) %>% 
  # preserve order of names
  mutate(outcome_names = parse_factor(outcome_names, levels = outcome))

@@

# print table
knitr::kable(sameday_results, 
  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits")
```

@@

### 7.b Age
*Jingyang to complete using function*
Color palette to use for the 3 age categories. c("#899DA4", "#35274A" "#F2300F").

##8. SABA same-day results by underlying respiratory diagnosis
*Jingyang to complete using function*
Color palette to use for the 4 age categories.
c("#E2D200" "#46ACC8" "#E58601" "#B40F20")

@@

Stratify cumulative effect by sex and age group.

@@

### Comments:
1. File structure needs to be cleaned up. There are way too many subdirectories.
2. Remove r code files that are not useful. There are too many files with similar code inside.
3. Variables in the casecross over datasets raise some questions.
4. Consider a cumulative effect that adds the preceding day and so on, i.e. 0, 0+1, 0+1+2, ...

>>>>

Packages used: tidyverse, survival.
```{r setup}
library(tidyverse) # general data wrangle
library(survival) # conditional logistic regression

# knitr options
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

@@

```{r smoke data import and wrangle}
# pm path
pm_path <- paste0("../../data/pm/2013-oregon_county_pm25.csv")
# county pm
county_pm_df <- read_csv(pm_path) %>% 
  mutate(county = stringr::str_replace(county, "[.]", ""))

@@

```{r health data import, message=F, warning=F}
# path for health dataframe
health_path <- "../../data/health/"
# importing time-stratified case-crossover dataframe
casecross_files <- list.files(path=health_path, pattern="*casecross*")

@@

```{r descriptive table}
# creating a row calculation function
table1_fun <- function(data_list){
  # output of total n
  total_n <- data_list %>% 
    filter(outcome == 1) %>%
    # adding in a filter to remove NAs
    filter(!is.na(geo_smk_pm_zip)) %>% 
    summarise(total_n = n())

@@

# apply custom function to generate descriptive table
table_vals <- outcome_df_list %>%
  # map to dataframe
  map_dfr(~table1_fun(.))

# descriptive table, bind outcome names and values
descriptive_table <- bind_cols(outcome, table_vals)

@@

```{r distributed lag}
# this approach will work that creates a list of dataframes with map then
# binds the lists together in one dataframe
dl_results <- outcome_df_list %>% 
  purrr::map(~distributed_lag(., pm_df=3,temp_df=2,cumulative=F)) %>% 
  plyr::rbind.fill() %>% 
  bind_cols(., data_frame(rep(outcome_names, each=8))) %>% 
  rename(outcome_names = `rep(outcome_names, each = 8)`) %>% 
  select(outcome_names, type, time, estimate, lower95, upper95) %>% 
  # preserve order of names
  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))

@@

```{r same day association stratified by sex}
# there is probably a better way to do this, but this will work.
sameday_results_f <- outcome_df_list %>% 
  # using gender because I forget which sex_ind is which
  # filter to females
  purrr::map(~filter(., gender == "F")) %>%  
  map_dfr(~same_day_fun(.)) %>% 
  bind_cols(., data_frame(outcome_names)) %>% 
  mutate(sex = as.factor("Female")) %>% 
  select(outcome_names, sex, estimates, lower95, upper95) %>% 
  # preserve order of names
  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))

sameday_results_m <- outcome_df_list %>% 
  # filter to males
  purrr::map(~filter(., gender == "M")) %>%  
  map_dfr(~same_day_fun(.)) %>% 
  bind_cols(., data_frame(outcome_names)) %>% 
  mutate(sex = as.factor("Male")) %>% 
  select(outcome_names, sex, estimates, lower95, upper95) %>% 
  # preserve order of names
  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))

@@

# print table
knitr::kable(sameday_results, 
  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits stratified by sex.")
```

@@

### 7.b Age

Stratified by age category.

```{r same day association stratified by age category}
# subset outcomes evaluated for under 15
under15_outcome <- outcome_names[1:7]

# subset dataframes that won't have convergence issues
sameday_results_under15 <- outcome_df_list[1:7] %>% 
  # filter to children
  purrr::map(~filter(., age_ind == 0)) %>%  
  map_dfr(~same_day_fun(.)) %>% 
  bind_cols(., data_frame(under15_outcome)) %>% 
  rename(outcome_names = under15_outcome) %>% 
  mutate(age_cat = as.factor("Under 15")) %>% 
  select(outcome_names, age_cat, estimates, lower95, upper95) %>% 
  # preserve order of names
  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))

# age category 1
sameday_results_15to65 <- outcome_df_list %>% 
  # filter to males
  purrr::map(~filter(., age_ind == "1")) %>%  
  map_dfr(~same_day_fun(.)) %>% 
  bind_cols(., data_frame(outcome_names)) %>% 
  mutate(age_cat = as.factor("15 to 65")) %>% 
  select(outcome_names, age_cat, estimates, lower95, upper95) %>% 
  # preserve order of names
  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))

# age category 2
sameday_results_over65 <- outcome_df_list %>% 
  purrr::map(~filter(., age_ind == "2")) %>%  
  map_dfr(~same_day_fun(.)) %>% 
  bind_cols(., data_frame(outcome_names)) %>% 
  mutate(age_cat = as.factor("Over 65")) %>% 
  select(outcome_names, age_cat, estimates, lower95, upper95) %>% 
  # preserve order of names
  mutate(outcome_names = parse_factor(outcome_names, levels = outcome_names))

# bind sex dataframes together
sameday_results <- rbind(sameday_results_under15, sameday_results_15to65,
                         sameday_results_over65)

# print table
knitr::kable(sameday_results, 
  caption = "Same day odds ratio and 95% CI of pharmacy/ED/urgent care visits stratified by age category.")
```

Stratified by age plot.
```{r age strata plot}
plot <- ggplot(data=sameday_results, 
               aes(x=age_cat, y = estimates, color=age_cat)) +
  scale_color_manual(name = "Age Category", 
    values = c("#899DA4", "#35274A", "#F2300F"),
    guide = guide_legend(direction="horizontal", title.position = "top",
                         title.hjust = 0.5)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  facet_wrap(~outcome_names, scales = "free_y") +
  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black")) +
  ylab(expression("Odds ratio: 10 ug/m^3 increase smoke PM"[2.5])) +
  xlab("Age Category")

Under 15 unstable for CVD; consider removing. Heart failure is interesting for the over 65 group. That may agree with what Anna Rappold has seen.

##8. SABA same-day results by underlying respiratory diagnosis

```{r saba underlying dx}
# read in the index 
saba_index <- read_csv("../../data/health/oregon_saba_index.csv") %>% 
    mutate(underlying_dx = as.factor(ifelse(index == "copd",
    stringr::str_to_upper(index),
    stringr::str_to_title(index)))) %>% 
    select(-index)

# index vector
underlying_dx <- unique(as.factor(saba_index$underlying_dx)) %>% 
  data_frame() %>% 
  rename(underlying_dx = ".") %>% 
  mutate(underlying_dx = ) %>% 
  # need to arrange since slice is going alphabetical
  arrange(underlying_dx)

# join index to saba
# extract saba dataframe from list
saba <- outcome_df_list[[1]] %>% 
  left_join(saba_index, by = "personkey")

# run function on each group
sameday_results <- saba %>% 
  split(.$underlying_dx) %>% 
  map_dfr(~same_day_fun(.)) %>% 
  bind_cols(., underlying_dx) %>% 
  select(4,1:3)

# table
knitr::kable(sameday_results, 
  caption = "Same day odds ratio and 95% CI for a SABA fill, stratified by underlying dx.")
```

*Jingyang, did I ask you to code "other" if a dx was missing? Maybe we should change other to "unknown or other" since I doubt we were able to link an underlying dx to everyone with a fill.*

Plot of SABA underlying Dx categories.

```{r plot saba underlying dx}
plot <- ggplot(sameday_results, 
   aes(x=underlying_dx, y = estimates, color = underlying_dx)) +
   scale_color_manual(name = "Underlying Diagnosis", 
    values = c("#0B775E","#46ACC8","#E58601","#B40F20"),
    guide = guide_legend(direction="horizontal", title.position = "top",
                         title.hjust = 0.5)) +
  geom_point() +
  geom_errorbar(aes(ymin=lower95, ymax=upper95), width = 0.2) +
  geom_hline(yintercept = 1, linetype = 2, colour = "#FF0000") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
   panel.grid.major = element_blank(),
   panel.grid.minor = element_blank(),
   axis.title.x = element_blank(),
   legend.position = "bottom") +
  ylab(expression("Odds Ratio: 10 ug/m^3 increase smoke PM"[2.5])) 
# plot
plot
```

@@

@@

### Issues/Concerns (in order of time they popped in my head):
1. File structure needs to be cleaned up. There are way too many subdirectories.
2. Remove r code files that are not useful. There are too many files with similar code inside.
3. Variables in the casecross over datasets raise some questions.
4. Consider a cumulative effect that adds the preceding day and so on, i.e. 0, 0+1, 0+1+2, ...
5. Plots need to have standard labels, ordering, scales, etc.
